title: "PATHOFigures_for_manu_FEB22"
---
author: "Kelsey Peterson"
date: "2/9/2022"
output: html_document
---

#ALL Packages needed for all the code below this chunk:


```{r, include = FALSE }
library(readr)
library(tidyverse)
library(dplyr)
library(ggrepel)
library(cowplot)
library(pixiedust)
library(kableExtra)
library(maps)
library(ggpubr)
library(ggExtra)
library(readxl)

```

---

 # Figure 1 - MAP
 
  Data and Code: Source: "Figure1_DimPatho.RMD"

```{r }
Dim2blatlong <- read_csv("Dim2b_latlon.csv")

more_info <- read_csv("2B_DIMENSIONS_data_compile.csv") %>%
  mutate(effective_precip = `30year_normal_ppt` - evapo_trans) %>%
  dplyr::select(effective_precip, Site_Name_Prairies)


Dim2blatlong <- full_join(Dim2blatlong , more_info) %>%
  filter(!is.na(Region))

my_map4 <- map_data("state") %>%
  filter(region %in% c("illinois", "kansas", "missouri"))

highlight_df <- Dim2blatlong %>%
  filter(Field_Type == "Common_Garden")
```

```{r}

pmain <- ggplot(data = Dim2blatlong,
                aes(x = lon, y = lat, color = effective_precip, fill = Field_Type,
                    show.legend = FALSE)) +
  geom_polygon(data = my_map4,
               aes(x = long, y = lat, group = group),
               color = "gray80",
               fill = "gray90",
               inherit.aes = FALSE) +
  geom_point(size = 5.5, aes(shape = Field_Type),alpha = .8) +
  geom_point( data =highlight_df, aes(x = lon, y = lat),
              color = "black", shape = 4, show.legend = FALSE,
              size = 8)+
  scale_fill_manual(values=c("blue", "black")) +
  scale_colour_gradient2(low = "#D55E00", mid = "#009E73",
                         high = "#0072B2",
                         midpoint = mean(Dim2blatlong$effective_precip,na.rm=TRUE))+
  scale_shape_manual(values=c(4, 16 ))+
  coord_fixed(1.3) +
  theme_classic() +
  labs(x = "Longitude", y = "Latitude")  +
  geom_text_repel(data = .%>% filter(Field_Type == "Common_Garden"), aes( label = RTlocs, colour = "black"),
                  show.legend = FALSE,
                  nudge_x = 0,
                  box.padding = 0.01,
                  point.padding = .5,
                  nudge_y = 1.5,
                  size = 3.5,
                  color = "Black",
                  check_overlap = TRUE) +
  geom_text(data = .%>% filter(Field_Type == "Prairie"),
                  aes( label = Site_letters, colour = "black"),
                  show.legend = FALSE,
                  nudge_x = 0,
                  nudge_y = 0,
                  box.padding = 0.0,
                  point.padding = 0,
                  size = 2.5,
                  color = "Black",
                  check_overlap = TRUE,
                  fontface="bold") +
  theme(legend.position = c(0.25, 0.85), legend.direction = "horizontal")+
  theme(legend.spacing.y = unit(1, "mm"),
        legend.background = element_blank(),
        legend.box.background = element_rect(colour = "black"))+
  labs(color = "Effective Precipitation")


pmain

```









---

#Figure 2 - Map of Disease scores in prairie remnants across years

CODE from : figure2_DimPatho.rmd
Original data: 2019and2020Prairie_data.txt -> Summaryleafrust, Summaryleafblotch, Summaryviral19, Summaryviral20 in the PrairieData_KP_april2021.Rmd


```{r, include=FALSE}
rm(list=ls(all=TRUE))
Summaryleafrust <- read_csv("Summaryleafrust")

Summaryleafrust$Region.f=ifelse(Summaryleafrust$Region.f== "C","Central", 
                                ifelse(Summaryleafrust$Region.f =="W","West",
                                       ifelse(Summaryleafrust$Region.f=="E","East", "NA")))

Summaryleafrust$Region.f<-factor(Summaryleafrust$Region.f,levels=c("West", "Central", "East"))

Summaryleafrust <- Summaryleafrust %>% separate(ComboName,
                c("Source_Prairie", "SP_Other_name"), " - ")

Summaryleafrust$Source_Prairie <- factor(Summaryleafrust$Source_Prairie,levels = c("W1 ", "W2 ", "W3 ", "W4 ",
                                                                                       "C1 ", "C2 ", "C3 ", "C4 ",
                                                                                       "E1 " , "E2 ", "E3 ", "E4 "))
```

```{r}
kpyanivoriginal_AVGPlotleafrust <- Summaryleafrust %>%
  mutate(num_reg = as.numeric(Region.f),
         x_for_p = num_reg + (as.numeric(Source_Prairie)%%4) / 5 -.3)  %>%
  arrange(factor(Source_Prairie, levels = c("W1", "W2", "W3", "W4",
                                  "C1", "C2", "C3", "C4",
                                  "E1", "E2", "E3", "E4")), x_for_p)  %>%
  ggplot(aes(x=as.numeric(Region.f), y=leaf_rust)) +
  facet_grid(~Year.f)+
  geom_bar(stat = "summary",
           fun.y = "mean",
           position = position_dodge2(width = 0.9, preserve = "single"),
           show.legend = FALSE,
           fill = "gray80",
          # fill = c("#D55E00", "#009E73", "#0072B2", "#D55E00", "#009E73", "#0072B2" ),
           alpha = .9,
           width = .5) +
  
      stat_summary(fun.data = mean_se,  
                 geom = "pointrange",
                 color = "black")+

  geom_point(aes( x = x_for_p,  
                  color = Region.f), 
                   
                  show_legend = FALSE,
                  size = 4,
             
             alpha = .6) +
    guides(color = FALSE) +
    
   #guides(color=guide_legend(title="Prairies")) +
  
    geom_text(aes(label = Source_Prairie, color = "Black"), 
             size = 2.3,
             check_overlap = FALSE,
             position = position_dodge2(width = .8, preserve = "single"),
             show.legend = FALSE,
             color = "Black",
             fontface="bold") + 
             #nudge_x = 0,) +
#geom_segment(aes(x =  x_for_p, xend =  x_for_p, y =  leaf_rust - ci,  yend = leaf_rust + ci))+
  labs(title ="", x="", y="Leaf Rust (% of total leaf area affected)") +
 
  facet_grid(~Year.f)  +
                theme_bw(base_size = 10) +
 scale_color_manual(values=c("#D55E00", "#009E73", "#0072B2")) + 
  #scale_color_manual(values=c("#ff704d", "#ff3300","#b32400", "#801a00","#80ff80", "#33ff33","#009900", "#004d00","#b3d9ff", "#4da6ff","#0066cc", "#00264d")) +
  scale_x_discrete(limits=c("West","Central","East")) 
                              
kpyanivoriginal_AVGPlotleafrust
 
```

```{r, include=FALSE}
Summaryleafblotch <- read_csv("Summaryleafblotch")

Summaryleafblotch$Region.f=ifelse(Summaryleafblotch$Region.f== "C","Central", 
                                ifelse(Summaryleafblotch$Region.f =="W","West",
                                       ifelse(Summaryleafblotch$Region.f=="E","East", "NA")))

Summaryleafblotch$Region.f<-factor(Summaryleafblotch$Region.f,levels=c("West", "Central", "East"))

Summaryleafblotch <- Summaryleafblotch %>% separate(ComboName,
                c("Source_Prairie", "SP_Other_name"), " - ")



Summaryleafblotch_excludecrazies <- subset(Summaryleafblotch, N > 13)

Summaryleafblotch_nowest<- Summaryleafblotch %>% filter(Region.f != "West")


Summaryleafblotch_excludecrazies_andwest <- subset(Summaryleafblotch, Leaf_blotch < 5)

Summaryleafblotch$Source_Prairie <- factor(Summaryleafblotch$Source_Prairie,levels = c("W1 ", "W2 ", "W3 ", "W4 ",
                                                                                       "C1 ", "C2 ", "C3 ", "C4 ",
                                                                                       "E1 " , "E2 ", "E3 ", "E4 "))


```

```{r}
kpyanivoriginal_AVGPlotleafblotch <- Summaryleafblotch %>%
  mutate(num_reg = as.numeric(Region.f),
         x_for_p = num_reg + (as.numeric(Source_Prairie)%%4) / 5 -.3)  %>%
  arrange(factor(Source_Prairie, levels = c("W1", "W2", "W3", "W4",
                                  "C1", "C2", "C3", "C4",
                                  "E1", "E2", "E3", "E4")), x_for_p)  %>%
  ggplot(aes(x=as.numeric(Region.f), y=Leaf_blotch)) +
  facet_grid(~Year.f)+
  geom_bar(stat = "summary",
           fun.y = "mean",
           position = position_dodge2(width = 0.9, preserve = "single"),
           show.legend = FALSE,
           fill = "gray80",
          # fill = c("#D55E00", "#009E73", "#0072B2", "#D55E00", "#009E73", "#0072B2" ),
           alpha = .9,
           width = .5) +
      stat_summary(fun.data = mean_se,  
                 geom = "pointrange",
                 color = "black")+

  geom_point(aes( x = x_for_p,  
                  color = Region.f), 
                   
                  show_legend = FALSE,
                  size = 4,
             
             alpha = .6) +
    guides(color = FALSE) +
    
   #guides(color=guide_legend(title="Prairies")) +
  
    geom_text(aes(label = Source_Prairie, color = "Black"), 
             size = 2.3,
             check_overlap = FALSE,
             position = position_dodge2(width = .8, preserve = "single"),
             show.legend = FALSE,
             color = "Black",
             fontface="bold") + 
             #nudge_x = 0,) +
#geom_segment(aes(x =  x_for_p, xend =  x_for_p, y =  leaf_rust - ci,  yend = leaf_rust + ci))+
  labs(title ="", x="", y="Leaf Blotch (% of total leaf area affected)") +
 
  facet_grid(~Year.f)  +
                theme_bw(base_size = 10) +
 scale_color_manual(values=c("#D55E00", "#009E73", "#0072B2")) + 
  #scale_color_manual(values=c("#ff704d", "#ff3300","#b32400", "#801a00","#80ff80", "#33ff33","#009900", "#004d00","#b3d9ff", "#4da6ff","#0066cc", "#00264d")) +
  scale_x_discrete(limits=c("West","Central","East")) 
                              
kpyanivoriginal_AVGPlotleafblotch
 
```

```{r, include=FALSE}

Summaryviral19 <- read_csv("Summaryviral19")

Summaryviral19$Region.f=ifelse(Summaryviral19$Region.f== "C","Central", 
                                ifelse(Summaryviral19$Region.f =="W","West",
                                       ifelse(Summaryviral19$Region.f=="E","East", "NA")))

Summaryviral19$Region.f<-factor(Summaryviral19$Region.f,levels=c("West", "Central", "East"))

Summaryviral19 <- Summaryviral19 %>% separate(ComboName,
                c("Source_Prairie", "SP_Other_name"), " - ")

Summaryviral19<- Summaryviral19 %>% filter(Year.f == "2019")

Summaryviral19$Source_Prairie <- factor(Summaryviral19$Source_Prairie,levels = c("W1 ", "W2 ", "W3 ", "W4 ",
                                                                                       "C1 ", "C2 ", "C3 ", "C4 ",
                                                                                       "E1 " , "E2 ", "E3 ", "E4 "))

```



```{r}
Summaryviral20 <- read_csv("Summaryviral20")

Summaryviral20$Region.f=ifelse(Summaryviral20$Region.f== "C","Central", 
                                ifelse(Summaryviral20$Region.f =="W","West",
                                       ifelse(Summaryviral20$Region.f=="E","East", "NA")))

Summaryviral20$Region.f<-factor(Summaryviral20$Region.f,levels=c("West", "Central", "East"))

Summaryviral20 <- Summaryviral20 %>% separate(ComboName,
                c("Source_Prairie", "SP_Other_name"), " - ")

Summaryviral20<- Summaryviral20 %>% filter(Year.f == "2020")

Summaryviral20$Source_Prairie <- factor(Summaryviral20$Source_Prairie,levels = c("W1 ", "W2 ", "W3 ", "W4 ",
                                                                                       "C1 ", "C2 ", "C3 ", "C4 ",
                                                                                       "E1 " , "E2 ", "E3 ", "E4 "))
```


`
```{r}
SummaryViral19 <- Summaryviral19 %>%
  mutate(viral_percent = as.numeric(viral_percent),
         num_reg = as.numeric(Region.f),
         x_for_p = num_reg + (as.numeric(Source_Prairie)%%4) / 5 -.3)  %>%
  arrange(factor(Source_Prairie, levels = c("W1", "W2", "W3", "W4",
                                  "C1", "C2", "C3", "C4",
                                  "E1", "E2", "E3", "E4")), x_for_p)

```
  
```{r}
kpyanivoriginal_AVGPlotleafvirus19 <- SummaryViral19 %>%
  ggplot(aes(x=as.numeric(Region.f), y=viral_percent)) +
  facet_grid(~Year.f)+
  geom_bar(stat = "summary",
           fun.y = "mean",
           position = position_dodge2(width = 0.9, preserve = "single"),
           show.legend = FALSE,
           fill = "gray80",
          # fill = c("#D55E00", "#009E73", "#0072B2", "#D55E00", "#009E73", "#0072B2" ),
           alpha = .9,
           width = .5) +
      stat_summary(fun.data = mean_se,  
                 geom = "pointrange",
                 color = "black")+

  geom_point(aes( x = x_for_p,  
                  color = Region.f), 
                   
                  show_legend = FALSE,
                  size = 4,
             
             alpha = .6) +
    guides(color = FALSE) +
    
   #guides(color=guide_legend(title="Prairies")) +
  
   geom_text(aes(label = Source_Prairie, color = "Black"), 
             size = 2.3,
             check_overlap = FALSE,
             position = position_dodge2(width = .8, preserve = "single"),
             show.legend = FALSE,
             color = "Black",
             fontface="bold") + 
             #nudge_x = 0,) +
#geom_segment(aes(x =  x_for_p, xend =  x_for_p, y =  leaf_rust - ci,  yend = leaf_rust + ci))+
  labs(title ="", x="Prairie Region", y="Silphium Clear Vein (% severity)") +
 
  facet_grid(~Year.f)  +
                theme_bw(base_size = 10) +
 scale_color_manual(values=c("#D55E00", "#009E73", "#0072B2")) + 
  #scale_color_manual(values=c("#ff704d", "#ff3300","#b32400", "#801a00","#80ff80", "#33ff33","#009900", "#004d00","#b3d9ff", "#4da6ff","#0066cc", "#00264d")) +
  scale_x_discrete(limits=c("West","Central","East")) 
                              
kpyanivoriginal_AVGPlotleafvirus19
 
```

```{r}
Summaryviral20 <- Summaryviral20 %>%
  mutate(viral_symptoms = as.numeric(viral_symptoms),
         num_reg = as.numeric(Region.f),
         x_for_p = num_reg + (as.numeric(Source_Prairie)%%4) / 5 -.3)  %>%
  arrange(factor(Source_Prairie, levels = c("W1", "W2", "W3", "W4",
                                  "C1", "C2", "C3", "C4",
                                  "E1", "E2", "E3", "E4")), x_for_p)  



kpyanivoriginal_AVGPlotleafvirus20 <-  Summaryviral20 %>%
  ggplot(aes(x=as.numeric(Region.f), y=viral_symptoms)) +
  facet_grid(~Year.f)+
  geom_bar(stat = "summary",
           fun.y = "mean",
     
           show.legend = FALSE,
           fill = "gray80",
          
           alpha = .9,
           width = .5) +
      stat_summary(fun.data = mean_se,  
                 geom = "pointrange",
                 color = "black")+

  geom_point(aes( x = x_for_p,  
                  color = Region.f), 
                    
                  show_legend = FALSE,
                  size = 4,
             
             alpha = .6) +
    guides(color = FALSE) +
  
   geom_text(aes(label = Source_Prairie, color = "Black"), 
             size = 2.3,
             check_overlap = FALSE,
             position = position_dodge2(width = .8, preserve = "single"),
             show.legend = FALSE,
             color = "Black",
             fontface="bold") + 
             

  labs(title ="", x="Prairie Region", y="Silphium Clear Vein (Scaled Severity)") +
 
  facet_grid(~Year.f)  +
                theme_bw(base_size = 10) +
 scale_color_manual(values=c("#D55E00", "#009E73", "#0072B2")) + 
  scale_x_discrete(limits=c("West","Central","East")) 
   
                              
kpyanivoriginal_AVGPlotleafvirus20
 
```

NEW VIRAL DATA PANEL
```{r}

viral_panel2 <- plot_grid(kpyanivoriginal_AVGPlotleafvirus19, 
                          kpyanivoriginal_AVGPlotleafvirus20, 
                          
                                   ncol = 2)
viral_panel2
```


use gg arrange
```{r, fig.height= 9, fig.width=6}

bottom_row <- plot_grid(kpyanivoriginal_AVGPlotleafvirus19,
          kpyanivoriginal_AVGPlotleafvirus20, labels = c('C', 'D'), label_size = 8)


plot_grid(kpyanivoriginal_AVGPlotleafrust,
          kpyanivoriginal_AVGPlotleafblotch,
          bottom_row,
          nrow = 3, ncol = 1,
          #rel_widths = c(2.3, 1.5, 1.5),
          labels = c('A', 'B', ' ' ),
          label_size = 8)

```



---

#Figure 3: DVT Index of germplasm 'performance' across all the sites

Data and code from : Figure3_Dim_Patho.rmd

```{r}
rm(list=ls(all=TRUE))
DVT_index <- read_csv("DVT_index_work_kp_version.csv")

DVTindex <- DVT_index %>% group_by (source) %>%
  mutate( index = sum(score))

```


```{r, fig.height= 3, fig.width=4}

Mainindex <- DVTindex %>% 
  filter(disease =="blotch", destination == "West") %>%
#  mutate(source_seed = str_remove(source, "[:digit:]"))
  mutate(source_seed = factor(case_when(source %in% paste("C",1:4,sep ="")~ "Central",
                    source %in% paste("W",1:4,sep ="")~ "West",
                    source %in% paste("E",1:4,sep ="")~ "East"),
         levels = c("West","Central","East" ) ))%>%
  
  ggplot(aes(x = source_seed, y = index))+
  stat_summary(aes(group = source_seed), geom = "bar",
               color = "lightgrey",
               alpha = .3)+
  stat_summary(aes(group = source_seed,fill = source_seed), geom = "pointrange")+
  geom_text(aes(label = source),
             position = position_dodge2(width = 0.9, preserve = "single"),
            size = 2.5,
            fontface="bold")+
  geom_hline(yintercept = 0) + 
  

    guides(color = FALSE, fill = FALSE) +
  labs(title ="", x="Prairie Region", y="Overall Index Score") +
                theme_bw() +
 scale_color_manual(values=c("#D55E00", "#009E73", "#0072B2")) +
 scale_fill_manual(values=c("#D55E00", "#009E73", "#0072B2")) + 
 geom_jitter(aes(col=source_seed), 
             alpha = .5, 
             show.legend = FALSE, 
             message = FALSE, 
             position = position_dodge2(width = 0.9, preserve = "single"),
             size = 6) +
    ylim(-5, 5) + 
  theme_bw() +
  theme(axis.text=element_text(size=10))

```


```{r, fig.height= 3, fig.width=4}


Blotchindex <-  DVTindex %>% 
   filter(disease =="blotch") %>%

  mutate(source_seed = factor(case_when(source %in% paste("C",1:4,sep ="")~ "Central",
                    source %in% paste("W",1:4,sep ="")~ "West",
                    source %in% paste("E",1:4,sep ="")~ "East"),
         levels = c("West","Central","East" ) )) %>%
#  group_by(source_seed, source, score) %>%
  group_by(source_seed, source) %>%
  #summarise(avgscore = mean(score)) %>%
 # 
  ggplot(aes(x = source_seed, y = score))+
  stat_summary(aes(group = source_seed), geom = "bar",
               color = "lightgrey",
               alpha = .3)+
  stat_summary(aes(group = source_seed,fill = source_seed), geom = "pointrange")+
  geom_text(aes(label = source),
             position = position_dodge2(width = 0.9, preserve = "single"),
            size = 2.5,
            fontface="bold")+
  geom_hline(yintercept = 0) + 
  

    guides(color = FALSE, fill = FALSE) +
  labs(title ="", x="Prairie Region", y="Leaf Blotch") +
                theme_bw() +
 scale_color_manual(values=c("#D55E00", "#009E73", "#0072B2")) +
 scale_fill_manual(values=c("#D55E00", "#009E73", "#0072B2")) + 
 geom_jitter(aes(col=source_seed), 
             alpha = .5, 
             show.legend = FALSE, 
             message = FALSE, 
             position = position_dodge2(width = 0.9, preserve = "single"),
             size = 6) +
    ylim(-1.5, 1.5) + 
  theme_bw() +
  theme(axis.text=element_text(size=10))


```

```{r, fig.height= 3, fig.width=4}

Rustindex <-  DVTindex %>% 
   filter(disease =="rust") %>% 
#  mutate(source_seed = str_remove(source, "[:digit:]"))
  mutate(source_seed = factor(case_when(source %in% paste("C",1:4,sep ="")~ "Central",
                    source %in% paste("W",1:4,sep ="")~ "West",
                    source %in% paste("E",1:4,sep ="")~ "East"),
         levels = c("West","Central","East" ) )) %>%
  #group_by(source_seed, source, score) %>%
  group_by(source_seed, source) %>%
summarise(avgscore = mean(score)) %>%
  

  ggplot(aes(x = source_seed, y = avgscore))+
  stat_summary(aes(group = source_seed), geom = "bar",
               color = "lightgrey",
               alpha = .3)+
  stat_summary(aes(group = source_seed,fill = source_seed), geom = "pointrange")+
  geom_text(aes(label = source),
             position = position_dodge2(width = 0.9, preserve = "single"),
            size = 2.5,
            fontface="bold")+
  geom_hline(yintercept = 0) + 
  

    guides(color = FALSE, fill = FALSE) +
  labs(title ="", x="", y="Leaf Rust") +
                theme_bw() +
 scale_color_manual(values=c("#D55E00", "#009E73", "#0072B2")) +
 scale_fill_manual(values=c("#D55E00", "#009E73", "#0072B2")) + 
 geom_jitter(aes(col=source_seed), 
             alpha = .5, 
             show.legend = FALSE, 
             message = FALSE, 
             position = position_dodge2(width = 0.9, preserve = "single"),
             size = 6) +
    ylim(-1, 1) + 
  theme_bw() +
  theme(axis.text=element_text(size=10))


```


```{r, fig.height= 3, fig.width=4}


viralindex <-  DVTindex %>% 
   filter(disease =="VeinClearing") %>% 
#  mutate(source_seed = str_remove(source, "[:digit:]"))
  mutate(source_seed = factor(case_when(source %in% paste("C",1:4,sep ="")~ "Central",
                    source %in% paste("W",1:4,sep ="")~ "West",
                    source %in% paste("E",1:4,sep ="")~ "East"),
         levels = c("West","Central","East" ) )) %>%
  #group_by(source_seed, source, score) %>%
  group_by(source_seed, source) %>%
summarise(avgscore = mean(score)) %>%


  ggplot(aes(x = source_seed, y = score))+
  stat_summary(aes(group = source_seed), geom = "bar",
               color = "lightgrey",
               alpha = .3)+
    geom_text(aes(label = source),
             position = position_dodge2(width = 0.9, preserve = "single"),
            size = 2.5,
            fontface="bold")+
  stat_summary(aes(group = source_seed,fill = source_seed), geom = "pointrange")+

  geom_hline(yintercept = 0) + 
  

    guides(color = FALSE, fill = FALSE) +
  labs(title ="", x="", y="Clear Leaf Vein Symptoms") +
                theme_bw() +
 scale_color_manual(values=c("#D55E00", "#009E73", "#0072B2")) +
 scale_fill_manual(values=c("#D55E00", "#009E73", "#0072B2")) + 
 geom_jitter(aes(col=source_seed), 
             alpha = .5, 
             show.legend = FALSE, 
             message = FALSE, 
             position = position_dodge2(width = 0.9, preserve = "single"),
             size = 6) +
    ylim(-1.5, 1.5) + 
  theme_bw() +
  theme(axis.text=element_text(size=10))

Rustindex
Blotchindex
viralindex
Mainindex

```

```{r, fig.height=3, fig.width=4}

toprowindex <- plot_grid(Rustindex,
          Blotchindex,
          viralindex,
          labels = c('(B)', '(C)', '(D)'),
          nrow = 1)

plot_grid <- plot_grid(Mainindex,
          toprowindex,
          nrow = 2, ncol = 1,
          rel_heights = c(1, 1),
          labels = c("(A)", ""))

```


---

#Figure 4: line graphs of disease scores across the common gardens 
From: Figure6_DimPatho.rmd

source: Zscores_Mar2021.R

```{r, include=FALSE, error=FALSE}
rm(list=ls(all=TRUE))
SummaryleafBlotchZ <- read_csv("SummaryleafBlotchZ")

SummaryleafBlotchZ$original_region<-factor(SummaryleafBlotchZ$original_region,levels=c("West", "Central", "East"))

SummaryleafBlotchZ$Location.f=ifelse(SummaryleafBlotchZ$Location.f== "central","Central", 
                                ifelse(SummaryleafBlotchZ$Location.f =="west","West",
                                       ifelse(SummaryleafBlotchZ$Location.f=="east","East", "NA")))

SummaryleafBlotchZ$Location.f<-factor(SummaryleafBlotchZ$Location.f,levels=c("West", "Central", "East"))

SummaryleafrustZ <- read_csv("SummaryleafrustZ")

SummaryleafrustZ$original_region<-factor(SummaryleafrustZ$original_region,levels=c("West", "Central", "East"))

SummaryleafrustZ$Location.f=ifelse(SummaryleafrustZ$Location.f== "central","Central", 
                                ifelse(SummaryleafrustZ$Location.f =="west","West",
                                       ifelse(SummaryleafrustZ$Location.f=="east","East", "NA")))

SummaryleafrustZ$Location.f<-factor(SummaryleafrustZ$Location.f,levels=c("West", "Central", "East"))

```

source: Zscores_Mar2021_Part2_kp.R, Eric

ONLY 2020

```{r, include=FALSE, error=FALSE}

SummaryViralZ <- read_csv("SummaryViralZ")

SummaryViralZ$Location.f=ifelse(SummaryViralZ$Location.f== "central","Central", 
                                ifelse(SummaryViralZ$Location.f =="west","West",
                                       ifelse(SummaryViralZ$Location.f=="east","East", "NA")))


SummaryViralZ$Location.f<-factor(SummaryViralZ$Location.f,levels=c("West", "Central", "East"))
```


```{r}
rustnew_zscore <- ggplot(data = SummaryleafrustZ, 
                         aes(x= Location.f, y=LeafRustZScore,  color = original_region)) +

  labs(title ="Rust Z Score", x="", y="Standard Deviations", size = 11) +
  stat_summary(position = position_dodge(width = .02)) +
  stat_summary( geom = "line", 
                lty = 1,
                alpha = .4,
                aes( group = original_region), 
               position = position_dodge(width = .02)) + 
  geom_linerange(aes(ymin=LeafRustZScore-se, ymax=LeafRustZScore+se), 
                position = position_dodge(width = .02)) +
  theme_bw() +
  labs(fill = "Original Prairie")+
  facet_grid()+ 
  theme(legend.position= "none") +
        #legend.background = element_rect(
         #                                linetype="solid", 
          #                               colour = "Black"))+
  guides(col=guide_legend("Original Region")) + 
  
  scale_color_manual(values=c("#D55E00", "#009E73", "#0072B2"))+
    scale_fill_manual(values=c("#D55E00", "#009E73", "#0072B2"))+
  
  geom_hline(yintercept = 0, col = "Black", lty = 2)+
  coord_cartesian(xlim = c(1.30, 1.8))+

  ylim(-1, 1) +
  theme(axis.text=element_text(size=12))
#    geom_text_repel(data = .%>% filter(Location.f == "East"), 
 #                aes(label = original_region),
  #               #repel to right
   #           show.legend = FALSE,
    #          nudge_x = .1,
     #         size = 4.5,
      #        point.padding = 0,) +
   #annotate("text", x=2.2, y=.85, label= "Source Region", fontface = 2, size = 4.5) +
  


rustnew_zscore 

```


```{r}

blotchnew_zscore <- ggplot(data = SummaryleafBlotchZ, aes(x= Location.f, y=LeafBlotchZScore,  color =  original_region)) +
 # geom_bar(stat = "identity", position = position_dodge())+  
  #geom_point(position = position_jitterdodge(dodge.width = .25, 
    #                                        jitter.width = .1 ),
   #          alpha  = .4) + #less jitter
  labs(title ="Leaf Blotch Z Score", x="", y="Standard Deviations", size = 11) +
  stat_summary(position = position_dodge(width = .1)) +
  stat_summary( geom = "line", 
                lty = 1,
                alpha = .4,
                aes( group = original_region), 
               position = position_dodge(width = .1)) + 
  geom_linerange(aes(ymin=LeafBlotchZScore-se, ymax=LeafBlotchZScore+se), 
                position = position_dodge(width = .1), show.legend = FALSE, message = FALSE ) +
  theme_bw() +
  labs(fill = "Original Prairie")+
  facet_grid()+ 
  theme(legend.title = element_blank(), legend.position = "none")+
  
  scale_color_manual(values=c("#D55E00", "#009E73", "#0072B2"))+
    scale_fill_manual(values=c("#D55E00", "#009E73", "#0072B2"))+
  
  geom_hline(yintercept = 0, col = "Black", lty = 2)+
 coord_cartesian(xlim = c(1.25, 2.8))+
  ylim(-1, 1) +
  theme(axis.text=element_text(size=12))
      
#coord_cartesian(xlim = c(1.5, 2.5))

blotchnew_zscore 

```


```{r}

new_zscore_viral <- ggplot(data = SummaryViralZ, aes(x= Location.f, y=ViralZScore,  color =  original_region)) +
 # geom_bar(stat = "identity", position = position_dodge())+  
  #geom_point(position = position_jitterdodge(dodge.width = .25, 
    #                                        jitter.width = .1 ),
   #          alpha  = .4) + #less jitter
  labs(title ="Viral Z Score", x="Common Garden", y="Standard Deviations", size = 11) +
  stat_summary(position = position_dodge(width = .1)) +
  stat_summary( geom = "line", 
                lty = 1,
                alpha = .4,
                aes( group = original_region), 
               position = position_dodge(width = .1)) + 
  geom_linerange(aes(ymin=ViralZScore-se, ymax=ViralZScore+se), 
                position = position_dodge(width = .1), ) +
  theme_bw() +
  labs(fill = "Original Prairie")+
  facet_grid()+ 
  theme(legend.title = element_blank(), legend.position = "none")+
  
  scale_color_manual(values=c("#009E73", "#0072B2", "#D55E00"))+
    scale_fill_manual(values=c( "#009E73", "#0072B2", "#D55E00"))+
  
  geom_hline(yintercept = 0, col = "Black", lty = 2)+
 coord_cartesian(xlim = c(1.25, 2.8))+
  ylim(-1.2, 1.2) +
  theme(axis.text=element_text(size=12))

 
new_zscore_viral 
```


Z SCORE PANEL
```{r, fig.width=3, fig.height= 9}

nofacet_zscore_panel <- plot_grid(rustnew_zscore, blotchnew_zscore, new_zscore_viral , 
                                   labels = c('A', 'B', 'C'), 
                                   label_size = 11,
                                   #nrow = 1,
                                   ncol = 1,
                                   show.legend = FALSE)
nofacet_zscore_panel

```


AB PLOTS

From: Figure6_DimPatho.rmd

Leaf Rust - AB Plots - Source: Graphs_2b_Mar2020_kp.rmd
```{r, include=FALSE}

SummaryleafrustABEW  <- read_csv("SummaryleafrustABEW")

SummaryleafrustABEW$Location.f=ifelse(SummaryleafrustABEW$Location.f== "central","Central", 
                                ifelse(SummaryleafrustABEW$Location.f =="west","West",
                                       ifelse(SummaryleafrustABEW$Location.f=="east","East", "NA")))
SummaryleafrustABEW$Location.f<-factor(SummaryleafrustABEW$Location.f,levels=c("West", "Central", "East"))
SummaryleafrustABEW$source_prairie<-factor(SummaryleafrustABEW$source_prairie,levels=c("W1", "W2", "W3", "W4",
                                                                                       "C1", "C2", "C3", "C4",
                                                                                       "E1", "E2", "E3", "E4"))
SummaryleafrustABEW$source_prairie2 = SummaryleafrustABEW$source_prairie
SummaryleafrustABEW <- SummaryleafrustABEW %>% separate(source_prairie2,
                c("Source_Prairie_letter", "SP_Other_name"), sep = "(?<=[A-Za-z])(?=[0-9])")
SummaryleafrustABEW$Source_Prairie_letter<-factor(SummaryleafrustABEW$Source_Prairie_letter,levels=c("W", "C", "E"))

```

Blotch - AB Plots - Source: Graphs_2b_Mar2020_kp.rmd
```{r, include=FALSE}
SummaryblotchABPrairie <- read_csv("SummaryblotchABPrairie")

SummaryblotchABPrairie$Location.f=ifelse(SummaryblotchABPrairie$Location.f== "central","Central", 
                                ifelse(SummaryblotchABPrairie$Location.f =="west","West",
                                       ifelse(SummaryblotchABPrairie$Location.f=="east","East", "NA")))
SummaryblotchABPrairie$Location.f<-factor(SummaryblotchABPrairie$Location.f,levels=c("West", "Central", "East"))
SummaryblotchABPrairie$source_prairie<-factor(SummaryblotchABPrairie$source_prairie,levels=c("W1", "W2", "W3", "W4",
                                                                                       "C1", "C2", "C3", "C4",
                                                                                       "E1", "E2", "E3", "E4"))
SummaryblotchABPrairie$source_prairie2 = SummaryblotchABPrairie$source_prairie
SummaryblotchABPrairie <- SummaryblotchABPrairie %>% separate(source_prairie2,
                c("Source_Prairie_letter", "SP_Other_name"), sep = "(?<=[A-Za-z])(?=[0-9])")
SummaryblotchABPrairie$Source_Prairie_letter<-factor(SummaryblotchABPrairie$Source_Prairie_letter,levels=c("W", "C", "E"))

```

Viral - AB Plots - Source: Graphs_2b_Mar2020_kp.rmd
```{r, include=FALSE}

SummaryviralABPrairie2 <- read_csv("SummaryviralABPrairie2")

SummaryviralABPrairie2$Location.f=ifelse(SummaryviralABPrairie2$Location.f== "central","Central", 
                                ifelse(SummaryviralABPrairie2$Location.f =="west","West",
                                       ifelse(SummaryviralABPrairie2$Location.f=="east","East", "NA")))
SummaryviralABPrairie2$Location.f<-factor(SummaryviralABPrairie2$Location.f,levels=c("West", "Central", "East"))
SummaryviralABPrairie2$source_prairie<-factor(SummaryviralABPrairie2$source_prairie,levels=c("W1", "W2", "W3", "W4",
                                                                                       "C1", "C2", "C3", "C4",
                                                                                       "E1", "E2", "E3", "E4"))
SummaryviralABPrairie2$source_prairie2 = SummaryviralABPrairie2$source_prairie
SummaryviralABPrairie2 <- SummaryviralABPrairie2 %>% separate(source_prairie2,
                c("Source_Prairie_letter", "SP_Other_name"), sep = "(?<=[A-Za-z])(?=[0-9])")
SummaryviralABPrairie2$Source_Prairie_letter<-factor(SummaryviralABPrairie2$Source_Prairie_letter,levels=c("W", "C", "E"))
```


```{r, fig.height=3, fig.width=8}

newABPlotleafrustEWFACET <- ggplot(data = SummaryleafrustABEW, 
                                   aes(x=Location.f, 
                                       y=leaf_rust, 
                                       group = source_prairie , 
                                       colour = source_prairie))+ 
  
  geom_line( alpha = .9)+
  geom_point( show.legend = FALSE,
              size = 1)+

  labs(title ="", 
       x="",
       y="Leaf Rust (% of total leaf area affected)",
       size = 2)+  
  theme_bw() +
    
  stat_summary(aes(group=Source_Prairie_letter, color = Source_Prairie_letter), 
               fun.y=mean,
               geom="line", 
               linetype = "dotted",
               size = 1,
               fullrange=TRUE,
               show.legend = FALSE) +
  guides(color=guide_legend(title="Source Prairies")) +
  
   scale_color_manual(values=c(
       #stat green
     "#00ba38",  
     #prairies
     "springgreen4", "darkgreen","chartreuse 2", "olivedrab",
      #stat blue
     "#619cff",
     #prairies'
     "darkcyan", "darkblue", "cyan", "slateblue3",
    # stat red:
     "#f8766d",
     #prairies: 
     "red", "red4","darkorange2", "firebrick2")) + 
  facet_grid(~Source_Prairie_letter)+
  
   coord_cartesian(xlim = c(1.25, 1.80))+
    
  geom_text_repel(data = .%>% filter(Location.f == "East"), 
                 aes( label = source_prairie),
                 show.legend = FALSE,
              size = 2,
              nudge_x = .2,
              check_overlap = TRUE,
              color = "black",
              min.segment.length = 0) +
  
  theme(legend.position = "none")+
  theme(axis.text=element_text(size=12))+
    theme(strip.text.x = element_text(size=10, colour = "white"),
          strip.background = element_rect(colour="black", fill="grey30")) 

newABPlotleafrustEWFACET


```


```{r, fig.height=3, fig.width=8}

newABPlotleafblotchFACET <- ggplot(data = SummaryblotchABPrairie,
                                   aes(x=Location.f, 
                                       y=leaf_blotch, 
                                       group = source_prairie , 
                                       colour = source_prairie))+ 
  
  geom_line( alpha = .9)+
  geom_point( show.legend = FALSE,
              size = 1)+

  labs(title ="", 
       x="",
       y="Leaf Blotch (% of total leaf area affected)",
       size = 2)+  
  theme_bw() +
    
  stat_summary(aes(group=Source_Prairie_letter, color = Source_Prairie_letter), 
               fun.y=mean,
               geom="line", 
               linetype = "dotted",
               size = 1,
               fullrange=TRUE,
               show.legend = FALSE) +
  guides(color=guide_legend(title="Source Prairies")) +
  

   scale_color_manual(values=c(
   
       #stat green
     "#00ba38",  
     #prairies
     "springgreen4", "darkgreen","chartreuse 2", "olivedrab",
      #stat blue
     "#619cff",
     #prairies'
     "darkcyan", "darkblue", "cyan", "slateblue3",
    # stat red:
     "#f8766d",
     #prairies: 
     "red", "red4","darkorange2", "firebrick2")) + 
  facet_grid(~Source_Prairie_letter)+
  
  
   coord_cartesian(xlim = c(1.25, 2.80))+
    geom_text_repel(data = .%>% filter(Location.f == "East"), 
                 aes( label = source_prairie),
                 show.legend = FALSE,
              size = 2,
              nudge_x = .2,
              check_overlap = TRUE,
              color = "black",
              min.segment.length = 0) +
  
   theme(legend.position = "none")+
  theme(axis.text=element_text(size=12)) +
  theme(strip.text.x = element_text(size=10, colour = "white"),
          strip.background = element_rect(colour="black", fill="grey30"))

newABPlotleafblotchFACET


```

```{r, fig.height=3, fig.width=8}

newABPlotleafviralFACET <- ggplot(data = SummaryviralABPrairie2, 
                                  aes(x=Location.f, 
                                      y=viral_symptoms, 
                                      group = source_prairie , 
                                      colour = source_prairie))+ 
  
  geom_line( alpha = .9)+
  geom_point( show.legend = FALSE,
              size = 1)+

  labs(title ="", 
       x="Common Garden",
       y="Silphium Clear Vein (Scaled Severity)",
       size = 2)+  
  theme_bw() +
    
  stat_summary(aes(group=Source_Prairie_letter, color = Source_Prairie_letter), 
               fun.y=mean,
               geom="line", 
               linetype = "dotted",
               size = 1,
               fullrange=TRUE,
               show.legend = FALSE) +
  guides(color=guide_legend(title="Source Prairies")) +
  
   scale_color_manual(values=c(
       #stat green
     "#00ba38",  
     #prairies
     "springgreen4", "darkgreen","chartreuse 2", "olivedrab",
      #stat blue
     "#619cff",
     #prairies'
     "darkcyan", "darkblue", "cyan", "slateblue3",
    # stat red:
     "#f8766d",
     #prairies: 
     "red", "red4","darkorange2", "firebrick2")) + 
  facet_grid(~Source_Prairie_letter)+
  
   coord_cartesian(xlim = c(1.25, 2.80))+
    geom_text_repel(data = .%>% filter(Location.f == "East"), 
                 aes( label = source_prairie),
                 show.legend = FALSE,
              size = 2,
              nudge_x = .2,
              check_overlap = TRUE,
              color = "black",
              min.segment.length = 0) +
   theme(legend.position = "none") +
  theme(axis.text=element_text(size=12)) +
    theme(strip.text.x = element_text(size=10, colour = "white"),
          strip.background = element_rect(colour="black", fill="grey30"))

newABPlotleafviralFACET


```


```{r, fig.height=8, fig.width=11}

ABZ <- plot_grid(newABPlotleafrustEWFACET, rustnew_zscore,
          newABPlotleafblotchFACET, blotchnew_zscore,
          newABPlotleafviralFACET, new_zscore_viral,
          nrow = 3, ncol = 2,
          rel_widths = c(1.5, 1),
          labels = "AUTO")

ABZ
```


                              
                              
                            